### 2022-06-04 manuscript.Rmd
Changed Spacing
Update manuscript.Rmd
Added to Starting Values
Added Assessing Estimation Results

    Mixture model parameters are estimated through maximum likelihood (ML) estimation with expectation-maximization (EM) as a local optimizer. Imagine we are estimating two parameters, e.g. the class-specific means $\mu_c$ on a continuous indicator (ignoring the variance for now). The EM algorithm will attempt to find a combination of values for these two parameters that maximizes the likelihood ($LL$) of all observed data. In practice, instead of maximizing $LL$, often $-2*LL$ is minimized, as this offers computational advantages. We can think of this optimization problem as a three-dimensional landscape: The X and Y dimensions are determined by the class-specific means, so $X = \mu_1$ and $Y = \mu_2$ - and the Z-dimension is determined by $Z = -2*LL$. The optimizer must find the deepest "valley" in this landscape, which reflects the combination of $\mu_1$ and $\mu_2$ that maximizes the likelihood of the data.
The EM optimizer behaves somewhat like a marble, dropped in this landscape.
It is dropped at some random point in space, and will roll into the nearest valley.
The problem is that, once EM rolls into a valley, it will settle on the bottom of that valley (this is known as "convergence").
It cannot climb out again.
Thus, if their are multiple valleys, the risk is that the optimizer gets stuck in a shallower valley (a "local optimum"), and never discovers the deepest valley (the "global optimum", or best solution).
One solution to this problem is to drop many marbles at random places, compare their final $-2*LL$ values,
choose the solution with the lowest $-2*LL$, and make sure that several marbles replicated this solution.
This is the "random starts" approach.

@@

By default, `tidySEM` eploys this solution of deriving starting values using K-means clustering,
and identifying the global optimum solution using simulated annealing.
Once a solution has been found,
simulated annealing is followed up with a short run of the EM algorithm,
as EM inherently produces an asymptotic covariance matrix for the parameters that can be used to compute standard errors.
Note that these defaults can be manually overridden.

@@

The algorithm needs starting values for each parameter it wants to estimate. If it were provided a single starting value for each parameter, the  EM optimizer would be at risk of finding a local optimum instead of the global optimum. In other words, it would converge on a suboptimal solution and the resulting parameter estimates would be biased. To prevent this, we should give the algorithm a sufficient set of random starting values for each parameter in order to maximize our chances of finding the deepest valley that is the global optimum (Geiser, 2013).

@@

>>>>

    Mixture model parameters and model fit statistics are estimated by maximum likelihood (ML) estimation with the expectation-maximization (EM) algorithm as a local optimizer. Imagine we are estimating two parameters, e.g. the class-specific means $\mu_c$ on a continuous indicator (ignoring the variance for now). The EM algorithm will attempt to find a combination of values for these two parameters that maximizes the likelihood ($LL$) of all observed data. In practice, instead of maximizing $LL$, often $-2*LL$ is minimized, as this offers computational advantages. We can think of this optimization problem as a three-dimensional landscape: The X and Y dimensions are determined by the class-specific means, so $X = \mu_1$ and $Y = \mu_2$ - and the Z-dimension is determined by $Z = -2*LL$. The optimizer must find the deepest "valley" in this landscape, which reflects the combination of $\mu_1$ and $\mu_2$ that maximizes the likelihood of the data.
The EM optimizer behaves somewhat like a marble, dropped in this landscape.
It is dropped at some random point in space, and will roll into the nearest valley.
The problem is that, once EM rolls into a valley, it will settle on the bottom of that valley (this is known as "convergence").
It cannot climb out again.
Thus, if their are multiple valleys, the risk is that the optimizer gets stuck in a shallower valley (a "local optimum"), and never discovers the deepest valley (the "global optimum", or best solution).
One solution to this problem is to drop many marbles at random places, compare their final $-2*LL$ values,
choose the solution with the lowest $-2*LL$, and make sure that several marbles replicated this solution.
This is the "random starts" approach.

@@

By default, `tidySEM` employs this solution of deriving starting values using K-means clustering,
and identifying the global optimum solution using simulated annealing.
Once a solution has been found,
simulated annealing is followed up with a short run of the EM algorithm,
as EM inherently produces an asymptotic covariance matrix for the parameters that can be used to compute standard errors.
Note that these defaults can be manually overridden.

@@

The EM algorithm needs starting values for each parameter it wants to estimate. 
If it were provided a single starting value for each parameter, 
the EM optimizer would be at risk of finding a local optimum instead of the global optimum. 
In other words, it would converge on a suboptimal solution resulting in biased parameter and model fit estimates. 
To prevent this, we should give the algorithm a sufficiently large set of random starting values for each parameter in order to maximize our chances of finding the deepest valley that is the global optimum (McLachlan & Peel, 2000). Furthermore, we should ensure the number of initial stage iterations is large for optimization to be successful (Geiser, 2013).

Furthermore, the choice of the starting values can impact the speed of the algorithm's convergence or in case of having poor initial values the algorithm might diverge altogether (McLachlan & Peel, 2000).
The key is to use a large number of starting value sets. Geiser (2013) recommends at least 500 sets of initial values in the first optimization step for simple models, and a greater number of sets for more complex models. A large number of starting values ensures that we find the true maximum and estimate the model parameters and fit accurately.

@@

### Assessing Estimation Results

Each set of starting values will result in a model with a log likelihood value. Ideally, the largest log likelihood value will be replicated a large number of times (each time is generated by a different starting value set). Successful replications of the largest log likelihood value across different starting values instill confidence that the estimation was successful. The model with the highest replicated likelihood will be chosen as our best-fit. If the largest log likelihood value is not replicated many times, the model may be unidentified (Geiser, 2013).

[to add: Nylund-Gibson & Choi  p.22. about 3-10% and Geiser p. 237 -240]
